---
title: "Lab 5"
author: "Nora Quick"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(1908) 
x <- rnorm(n = 40, mean = 5, sd = 5) 
```

# Instructions

Complete the lab tutorial before completing this file.  Use the R Markdown version of this file to complete and submit your homework.  Items in **bold** require an answer. Make sure you change the author in the header to your own name. 

1.  All three test functions (`t.test()`, `binom.test()` and `prop.test()`) have an argument called `alternative`. **What does this argument do?** Use the help to find out.

Alternative is used to specify the alternative hypothesis (greater or less than).

2.  In the `t.test()` section of the lab, you tested the hypotheses:
    $$
    H_{0}: \mu = 3, \quad H_{A}: \mu \neq 3 
    $$
    
    **Edit the code below to test the hypotheses:**
    $$
    H_{0}: \mu = 5, \quad H_{A}: \mu > 5 
    $$
    
    ```{r, results = "hide"}
    t.test(x, mu = 5, conf.level = 0.95)
    ```


3. In the `prop.test()` section of the lab, you tested the hypotheses:
    
    $$ 
    H_{0}: p = 0.25, \quad H_{A}: p \neq 0.25 
    $$
    based on observing 20 successes in 100 trials.
    
    **Edit the code below to test the same hypotheses, but based 
    on 200 successes in 1000 trials:**
    
    ```{r, results = "hide"}
    prop.test(x = 200, n = 1000, p = 0.25, 
      conf.level = 0.95, correct = FALSE)
    ```

4.  In the `prop.test()` section of the lab, you tested the hypotheses:
    
    $$ 
    H_{0}: p = 0.25, \quad H_{A}: p \neq 0.25 
    $$
    based on observing 20 successes in 100 trials.
    
    **Edit the code below to test the same hypotheses, with the same data,  but using an exact Binomial test instead. Would you reach the same conclusion?**

    ```{r, results = "hide"}
    binom.test(x = 20, n = 100, p = 0.25, 
      conf.level = 0.95)
    ```

Yes, with a test run of both we can see that we get the same result.