---
title: "Data Manipulation"
author: Nora Quick
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Learn

```{r}
library(tidyverse)
dog_licenses <- readr::read_csv("https://github.com/merely-useful/novice-r/raw/master/data/nyc-dog-licenses.csv.gz") %>% drop_na(animal_name)
```

```{r}
#View(dog_licenses)
#dog_licenses
```

```{r}
#arrange(dog_licenses, animal_birth_month)
#arrange(dog_licenses, desc(animal_birth_month))
```

```{r}
#arrange(dog_licenses, license_issued_date)
```

### 4.3.2 Exercise:
```{r}
arrange(dog_licenses, animal_name)
```

It shows that it will provide puncuation first, then numbers, then letters. In other words, puncuation and numbers are given first priority.

```{r}
#select(dog_licenses, animal_name)
#select(dog_licenses, animal_name, breed_name)
```

```{r}
#dog_by_date <- arrange(dog_licenses, license_issued_date)
#select(dog_by_date, license_issued_date)
```

```{r}
#select(dog_licenses, starts_with("Animal"))
```

```{r}
#?dplyr::select
```

```{r}
#arrange(dog_licenses, license_issued_date)
#dog_licenses %>% arrange(license_issued_date)
```

```{r}
#dog_licenses %>%
#  arrange(license_issued_date) %>%
#  select(license_issued_date)
```

### 4.4.5 Exercise:
```{r}
#select(dog_licenses, animal_name, breed_name)
dog_licenses %>% select(animal_name, breed_name)
```

```{r}
#name_and_breed <- select(dog_licenses, animal_name, breed_name)
#arrange(name_and_breed, breed_name)
dog_licenses %>%
  select(animal_name, breed_name) %>%
  arrange(breed_name)
```

```{r}
#dog_licenses %>% filter(animal_name == "BRUNO")
```

```{r}
#dog_licenses %>% filter(license_issued_date == animal_birth_month)
```

```{r}
#dog_licenses %>% filter("animal_gender" == "M")
#dog_licenses %>% filter(animal_gender == "M")
```

### 4.4.7 Exercise:
```{r}
dog_licenses %>% filter(animal_name == "SPOCK")
dog_licenses %>% filter(animal_name == "PICARD")
dog_licenses %>% filter(animal_name == "JANEWAY")
dog_licenses %>% filter(animal_name == "HARRY")
```

```{r}
#dog_licenses %>% filter(animal_name == "BRUNO")
#dog_licenses %>% filter(borough == "Brooklyn")
#dog_licenses %>% filter((animal_name == "BRUNO") & (borough == "Brooklyn"))
```

```{r}
#dog_licenses %>% filter((animal_name == "BRUNO") | (animal_name == "BRUCE"))
```

```{r}
#dog_licenses %>% filter((animal_name == "BRUNO") | (animal_name == "BRUCE") | (animal_name == "BRADY"))
```

```{r}
#dog_licenses %>% filter(animal_name %in% c("BRUNO", "BRUCE", "BRADY"))
```

### 4.4.10 Exercise:
```{r}
start_of_2016 <- as.Date("2016-01-01")
endt_of_2016 <- as.Date("2016-12-31")

dog_licenses %>% filter(license_expired_date %in% c(start_of_2016, endt_of_2016))
```

```{r}
#dog_licenses %>% 
#  mutate(called_chase = animal_name == "CHASE") %>% 
#  select(animal_name, called_chase)
```

```{r}
#dog_licenses %>%
#  mutate(called_chase = animal_name == "CHASE") %>%
#  filter(called_chase)
```

```{r}
#dog_licenses %>%
#  mutate(called_chase = ifelse(animal_name == "CHASE", "called chase", "not called chase")) %>%
#  select(animal_name, called_chase)
```

```{r}
#dog_licenses %>%
#  mutate(is_chase = animal_name == "CHASE",called_chase = ifelse(is_chase, "called chase", "not called chase")) %>%
#  select(animal_name, is_chase, called_chase)
```

```{r}
#dog_licenses %>%
#  mutate(license_duration = license_expired_date - license_issued_date) %>%
#  select(license_duration)

#dog_licenses %>%
#  mutate(license_duration = license_expired_date - license_issued_date,
#    avg_duration = mean(license_duration)) %>%
#  select(license_duration, avg_duration)
```

### 4.5.3 Exercise:
```{r}
dog_licenses %>%
  mutate(name_length = stringr::str_length(animal_name)) %>%
  arrange(desc(name_length))
```

There are four dogs with 30 character names with the classified longest name of 	CARLYAPPLEWHITECRAWFORDCOLEMAN.


## Apply
```{r}
us_pop <- read_rds("http://data.cwick.co.nz/us-population.rds")
us_pop
```

### 1.

```{r}
us_pop <- arrange(us_pop, date)
```

### 2.

```{r}
ggplot(us_pop, aes(date, population)) +
  geom_line()
```

### 3.

```{r}
us_pop <- us_pop %>% 
  mutate(prev_population = lag(population))
```

This compares the "current" population to the previous population. In other words, it compares each rows population to the population of the row above.

### 4. 
```{r}
us_pop <- us_pop %>%
  mutate(change_rate = (population/prev_population))
us_pop
```

### 5. 
```{r}
ggplot(us_pop, aes(date, change_rate)) +
  geom_line()
  #geom_point()
```

### 6. 
```{r}
us_pop <- us_pop %>% filter(year %in% ("1901":"1989"))

ggplot(us_pop, aes(date, change_rate)) +
  geom_line()
```

### 7. 
```{r}
us_pop <- us_pop %>%
  mutate(log10_population = log10(population))

ggplot(us_pop, aes(date, log10_population)) +
  geom_line()
```

The population has steadily increased over the years. 

### 8. 
```{r}
us_pop %>% 
  ggplot(aes(date, population)) +
    geom_line() +
    scale_y_log10(labels = scales::label_comma())
```

They both have a similar trend in increase rate. I find the second one easier to understand because the I find the populatoin number easier to understand. 


