---
title: "Lab 6"
author: "Nora Quick"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(1908) 
n <- 40
x <- rnorm(n = n, mean = 5, sd = 5) 
y <- x + 5 + rnorm(n)
z <- sample(x, size = 45, replace = TRUE) + 1 + rnorm(45)
```

# Instructions

Complete the lab tutorial before completing this file.  Use the R Markdown version of this file to complete and submit your homework.  Items in **bold** require an answer. Make sure you change the author in the header to your own name. 

1\.  Consider the code and output below:

```{r}
t.test(x, y, paird = TRUE)
```

a) **Has a paired or two-sample t-test been conducted?** 
A two-sample t-test has been conducted.

b) **What test do you think the user wanted, and can you see their mistake?**
They wanted to conduct a paired t-test, however, they forgot the 'e' in "paired = True".


2\. Consider the code and output below:

```{r}
t.test(x, z)
```

**Which of the following is the correct interpretation of the confidence interval?**
 
* With 95\% confidence, the population mean of `x` is between 0.14 and 3.94
units greater than the population mean of `z`.

* With 95\% confidence, the population mean of `x` is between 3.94 and 0.14
units less than the population mean of `z`.

* With 95\% confidence, the population mean of `x` is between 3.94 units less and
0.14 units greater than the population mean of `z`.

* With 95\% confidence, the population mean of `x` is between 0.14 units less 
and 3.94 units greater than the population mean of `z`.


The correct interpretation of the confidence interval is "With 95\% confidence, the population mean of `x` is between 3.94 units less and
0.14 units greater than the population mean of `z`."
