---
title: "Lab Assignment 2"
author: "Nora Quick"
output:
  pdf_document: default
  html_notebook: default
---

```{r setup, include=FALSE, message=FALSE}
library(magrittr)
library(vcdExtra)
library(tidyverse)
```
## Questions: 

### Q1.How consistent are these data with Mendel's hypothesis of independence?  Notice that in this particular case, we're not interested in a *departure* from independence, but rather a *confirmation* of independence. Is a Chi-squared test appropriate here? Why or why not? 

```{r}
tab1 <- matrix(c(315, 108, 101, 32),2,2,byrow=TRUE)
mosaic(tab1)

sieve(tab1, shade = TRUE)
```

Based on the two graphs above I believe that the data is approximately independent.  

Because the Chi-Squared test is used to test and no necessarily confirm independence (what we want to do here) I would say that it is not needed here.  


### Q2.

```{r}
norm <- rnorm(1000)
chi <- rchisq(1000, df = 1)
```

```{r}
sqnorm <- norm * norm
```

```{r}
dnorm <- density(sqnorm)
dchi <- density(chi)
```


#### (a) Construct separate histograms/density plots of the raw chi-squared values and the *squares* of the standard normal values. What do you observe?

```{r}
hist(sqnorm)
hist(chi)
```

```{r}
plot(dnorm)
plot(dchi)
```

I am very shocked to observe that in both cases of the histogram and the density are very similar. 

#### (b) Recall (or obtain from `qnorm(0.975)` that the *2-sided 95%* critical values for the standard normal distribution are ± 1.96. Use `qchisq()` to compare the *square* of these 2-sided 95% critical values from the standard normal to the *one-sided* 95% critical value for the Chi-squared(1).  What does this suggest about tests based on these two reference distributions?

```{r}
qnorm(0.975)
```

```{r}
sqrt(qchisq(chi, df = 1))
```





