---
title: "Homework Set 1"
data: "`r Sys.Date()`"
output: word_document
---

The goal of this first homework set is to help you

* set up R and R Studio, 
* set up the data collection to be used for this class, 
* run simple computations and simulations using R, and 
* learn to use R Markdown to create a word document with embedded R code and output.

## Task 1. Install R and R studio
Install `R` and `R Studio`. Follow instructions at

https://www.r-project.org/

and

https://www.rstudio.com/

## Task 2. Download the class data set and load one data set into R
1. Download the class data set (data.zip). Unzip it. Keep the folder name as `data`.
2. Create a folder `R` so that the folders `R` and `data` are at the same level.
3. Save this Rmd file into the `R` folder and open the Rmd file from the `R`
folder: this should automatically launch R Studio.

## Task 3. Compute normal probabilities.
1. Suppose $Z \sim N(0, 1)$ (standard normal, use R function `pnorm` to compute 
* $Pr(Z>2)$
* $Pr(-1 < Z \leq 1)$
* $Pr(0 < Z < 3)$

```{r}
## To compute probability Pr(Z>2)
1 - pnorm(2)
```

```{r}
## To compute probability Pr(-1<Z\leq1)
(1 - pnorm(1)) - pnorm(-1)
```

```{r}
## To compute probability Pr(0<Z<3)
pnorm(3) - pnorm(0)
```


2. Suppose $X \sim N(\mu=2, \sigma^2=5)$, use R function `pnorm` to compute
* $Pr(X > 5)$
* $Pr(X < 0)$
* $Pr(0 < X \leq 2)$

```{r}
## To compute Pr(X<0)
pnorm(0, mean = 2, sd = sqrt(5))
```

```{r}
## To compute Pr(X>5)
pnorm(5, mean = 2, sd = sqrt(5), lower.tail = FALSE)
```

```{r}
## To compute Pr(0<Z<\leq2)
pnorm(2, mean = 2, sd = sqrt(5)) - pnorm(0, mean = 2, sd = sqrt(5))
```


## Task 4. Simulation for Central Limit Theorem

```{r}
n_sim = 10000;

mean = 0.5;
sd = 0.5;

for (n in c(5, 10, 50, 100, 500, 1000)) {
  m = numeric(n_sim); 
  z = numeric(n_sim); 
  
  for (i in 1:n_sim) {
    
    y = rbinom(n, 1, p=0.5);

    m[i] = mean(y);

    z[i] = sqrt(n)* (m[i] - mean)/sd
    
    ## Or use the sample standard deviation
    ## z[i] = sqrt(n)* (mean(y) - mu)/sd(y);
  }

  hist(z, probability =TRUE, main=sprintf("n= %d", n), nclass=30);

  x = seq(-3, 3, length=100);
  lines(x, dnorm(x));

  plot(ecdf(z), col="magenta", main="Emperical CDF of z (magenta) to the normal cdf (cyan)");
  lines(x, pnorm(x), type="l", col="darkcyan");
}
```

## Task 5. Knit the R markdown file

Once all your R codes work correctly, "Knit" the markdown file to create the
word document. Verify that you have completed all the problems and then turn in
the completed word document through Canvas.

