---
title: "Module 3 Lab Submission"
author: "Nora Quick"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(broom)
```

Consider the `cars` data, which contains cars speed in MPH and stopping distance in feet. Load the data with `data("cars")`.

```{r, echo = TRUE}
data("cars")
qplot(speed, dist, data = cars)
```

* Fit a simple linear model with `dist` as the response and `speed` as the explanatory variable.

```{r}
fit <- lm(dist ~ speed, data = cars)
summary(fit)
```


* Create two diagnostic plots using the residuals, one with `speed` on the x-axis, and the other with the fitted values from the model. Do the plots look good: do these data seem to satisfy the assumptions for a linear regression model?

```{r}
f <- augment(fit)

qplot(speed, .resid, data = f)
qplot(.fitted, .resid, data = f)
```


* Use `predict()` to get the confidence and prediction intervals using the following new data.

```{r}
new <- data.frame(speed = c(6, 10.5, 14.7, 18.3, 21))

predict(fit, newdata = new, interval = "confidence")
```

Now note that there are many speeds for which there were multiple observations at that speed.  This means we can perform a lack-of-fit test on this data.  

* Fit a separate means model using `lm()` and `factor()` to treat `speed` as a categorical variable.

```{r}
lm_speed <- lm(dist ~ speed, data = cars)
lm_speed

fac_speed <- lm(dist ~ factor(speed), data = cars)
fac_speed
```

* Compare the separate means model to the simple linear regression model using the `anova()` function. 

```{r}
anova(lm_speed, fac_speed)
```



