---
title: "Module 5 Lab Submission"
author: "Nora Quick"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(broom)
library(ggplot2)
library(MASS)
library(Sleuth3)
```

Again we consider the Brain size data in the data set `case0902` from the `Sleuth3` library. You can read more about this data set by viewing the help file:

```{r echo=TRUE}
help(case0902)
head(case0902)
```

In Lecture 2 of this Module, we considered modeling all four variables on the log scale, and we compared two models: a full model with log-Body, log-Gestation, and log-Litter as predictors, and a reduced model with only log-Body as a predictor.

1. **Fit the two models described above (and as shown on slide 15 of Lecture 2).**

```{r echo=TRUE}
full_model <- lm(log(Brain) ~ log(Body) + log(Gestation) + log(Litter), data = case0902)

reduced_model <- lm(log(Brain) ~ log(Body), data = case0902)
```

Now suppose we have four new species, with the following values of Body, Gestation, and Litter:

| Species | Body | Gestation | Litter |
|:-------:|:----:|:---------:|:------:|
| A       | 1    | 50        | 1      |
| B       | 1    | 50        | 2      |
| C       | 1    | 150       | 1      |
| D       | 50   | 50        | 1      |


2. **Create a new data frame with these values, and use the `predict()` function to predict the log-brain size for these new species.**

```{r echo=TRUE}
same_gest <- data.frame(
  Species = rep(c('A', 'B', 'C', 'D'), 4),
  Body = rep(c(1, 1, 1, 50), each = 4),
  Gestation = rep(c(50, 50, 150, 50), each = 4),
  Litter = rep(c(1, 2, 1, 1), each = 4)
)

# same_gest
```

```{r echo=TRUE}
cbind(same_gest, pred_log_energy = predict(full_model, newdata = same_gest))
```

```{r echo=TRUE}
cbind(same_gest, pred_log_energy = predict(reduced_model, newdata = same_gest))
```


3. **Obtain confidence intervals for the mean log-brain size of the new species above, using the `interval = "confidence"` argument in the `predict()` function.**

```{r}
ci_full_model <- predict(full_model, newdata = same_gest, interval = "confidence")
ci_full_model
```

```{r}
ci_reduced_model <- predict(reduced_model, newdata = same_gest, interval = "confidence")
ci_reduced_model
```


4. **Obtain prediction intervals for the log-brain size of the new species above, using the `interval = "prediction"` argument in the `predict()` function.**

```{r}
pi_full_model <- predict(full_model, newdata = same_gest, interval = "prediction")
pi_full_model
```

```{r}
pi_reduced_model <- predict(reduced_model, newdata = same_gest, interval = "prediction")
pi_reduced_model
```


5. **Compare the reduced model to the full model using the AIC and BIC measures.**

```{r}
AIC_comp <- AIC(full_model, reduced_model)
AIC_comp
```
```{r}
BIC_comp <- BIC(full_model, reduced_model)
BIC_comp
```
