---
title: "ST 518 - HW2"
author: Nora Quick
output: pdf_document
---

# R Questions

```{r}
# left
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

# right
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

```{r}
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))
```


##2. Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.
This will create a plot with displ on the x-axis and hwy on the y-axis with colour bd on drv. It will add both points and a smooth line.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```
My prediction was for the most part accurate.

##3. What does show.legend = FALSE do? What happens if you remove it?
show.legend = FALSE removed the legend from the right side. IF you remove it, it will add the legend back.

##4. What does the se argument to geom_smooth() do?
It adds a grey boarder for confindence interval around the smooth lines. 

##5. Will these two graphs look different? Why/why not?
No, they will not look different because they have the same information in different formats. 

##6. 
















